Learning by Teaching a Robot: the Case of Handwriting

### 2015-07-07 main.tex
Working on the technical implementation

In order to develop a teachable agent that is appropriate for engaging a child
in the learning by teaching paradigm, we have established capabilities for the
robot to engage in handwriting and interactive turn-taking.

The {\sc nao} V4 humanoid robot, which has been purposely designed by Aldebaran
Robotics to look approachable~\cite{Gouaillier2008}, is used for this work. It
is a commercially affordable biped robot, 58cm tall, with 25 degrees of freedom,
two cameras, speech capabilities and the ability to autonomously execute a range
of tasks.

Precise control over what the robot is writing is necessary in the proposed
application of a handwriting. Because of the limited fine motor skills possible
with such an affordable robot, in addition to the absence of force feedback and
other technical necessities, we have configured the {\sc nao} to use simulated
handwriting with a synchronised tablet to achieve this level of control.

@@

@@

\resizebox{0.9\linewidth}{!}{%

\begin{tikzpicture}[
    >=latex,
    node distance=2cm,
    every edge/.style={draw, very thick},
    redarrow/.style={fill=Red, draw=Red},
    greenarrow/.style={fill=GreenYellow, draw=GreenYellow},
    yellowarrow/.style={fill=BurntOrange, draw=BurntOrange},
    cmpt/.style={draw, align=center, rounded corners, inner sep=5pt, font=\sf, fill=black!20},
    label/.style={midway, align=left, font=\scriptsize\sf, fill=white, above,opacity=0,text opacity=1}]

    \node at (0,0)[cmpt, fill=Cyan!50, text width=2.7cm] (tactile) {tactile interaction \\ manager};
    \node[cmpt, fill=Cyan!50, left=of tactile] (display) {display \\ manager};

    \node [cmpt, fill=Green!50, below=of tactile, text width=2.7cm] (shape) {shape position \\ manager};
    \node [cmpt, fill=Green!50, left=of shape] (learner) {letter \\ learner};
    \node [cmpt, fill=Green!50, left=of learner] (card) {card \\ detector};

    \node [cmpt, fill=YellowOrange!50, below=1.5cm of shape] (writer) {nao writer};
    \node [cmpt, fill=YellowOrange!50, left=of writer] (tablet) {tablet detector};
    \node [cmpt, fill=YellowOrange!50, left=of tablet] (camera) {nao camera};
    \node [cmpt, fill=YellowOrange!50, below=0.6cm of tablet] (pose) {robot state publisher};

    %%% Relations between components
    \path (tactile) edge [->] node[label] {clear \\ screen} (display);

    \path ($(shape.north west)!0.33!(shape.north east)$) edge [<-] node[label,left] {clear \\ screen} ($(tactile.south west)!0.33!(tactile.south east)$);
    \path ($(shape.north west)!0.66!(shape.north east)$) edge [<-, greenarrow] node[label,right] {feedback \\ + location} ($(tactile.south west)!0.66!(tactile.south east)$);

    \path (shape) edge [->, redarrow] node[label, left, align=left] {letter \\to write \\ + time \\ + position} (display);
    \path (shape) edge [->,redarrow] node[label, right] {letter to write \\ + time \\ + position in\\tablet frame} (writer);

    \path ($(shape.north west)!0.66!(shape.south west)$) edge [->, greenarrow] node[label,below] {feedback \\ + letter} ($(learner.north east)!0.66!(learner.south east)$);
    \path ($(shape.north west)!0.33!(shape.south west)$) edge [<-, redarrow] node[label] {letter to write \\ + time} ($(learner.north east)!0.33!(learner.south east)$);

    \path (card) edge [->, yellowarrow] node[label] {word to learn} (learner);
    \path (camera) edge [->, yellowarrow] node[label, right] {camera image} (card);
    \path (camera) edge [->, redarrow] node[label] {camera image} (tablet);
    \path (tablet) edge [->, redarrow] node[label] {tablet pose} (writer);
    \path (pose) edge [->, redarrow] node[label, right] {head pose} (tablet);

    \path (-8, 0) edge [->, redarrow] node[label] {Writing letters} (-7, 0);
    \path (-8, -0.6) edge [->, greenarrow] node[label] {Responding to feedback} (-7, -0.6);
    \path (-8, -1.2) edge [->, yellowarrow] node[label] {Detecting card input} (-7, -1.2);

\caption{Overview of the system. Components in the top row run on the tablet, 
those in the middle row on the central controller, and those in the bottom row 
on the robot.}

>>>>

Our system is embodied through an Aldebaran's {\sc nao} (V4 or V5, depending on the studies)
humanoid robot. This choice is motivated by its approachable
design~\cite{Gouaillier2008}, its size (58cm) and inherently safe structure (lightweight
plastic) (making it suitable for close interaction with children), its low price
(making it closer to what school may afford in the coming years) and finally its
ease of deployment on the field.

{\sc nao} is a humanoid biped, has 25 degrees of freedom,
two cameras, speech capabilities and the ability to autonomously execute a range
of tasks. While it can walk, this capability is not used in this project, and
the robot remains in a crouching posture during the whole interaction.

Robotic handwriting requires precise closed-loop control of the arm and hand
motion. Because of the limited fine motor skills possible
with such an affordable robot, in addition to the absence of force feedback and
other technical necessities, we have opted for \emph{simulated
handwriting}: the robot draws letters in the air, and the actual writing is
displayed on a synchronised tablet.

@@

\resizebox{0.6\linewidth}{!}{%

\begin{tikzpicture}[
    >=latex,
    node distance=2cm,
    every edge/.style={draw, very thick},
    redarrow/.style={draw=Red},
    greenarrow/.style={draw=GreenYellow},
    yellowarrow/.style={draw=BurntOrange},
    cmpt/.style={draw, align=center, rounded corners, inner sep=5pt, font=\sf, fill=black!20},
    label/.style={midway, align=left, font=\scriptsize\sf, fill=white, above,opacity=0,text opacity=1}]

    \node at (0,0) (laptop) {\includegraphics[width=2cm]{laptop}};
    \node[below right=2 of laptop] (nao) {\includegraphics[width=2cm]{nao}};
    \node[below left=2 of laptop] (tablet) {\includegraphics[width=2cm]{tablet+stylus}};
    \node[above=2 of laptop] (selection) {\includegraphics[width=2cm]{selection_tablet}};

    \path (nao) edge [->,redarrow, bend left] node[label, auto] {robot state} (laptop);
    \path (laptop) edge [->,greenarrow, bend left] node[label, auto] {writing gesture} (nao);

    \path (tablet) edge [->,redarrow, bend left] node[label, auto] {demonstrations,\\turn taking} (laptop);
    \path (laptop) edge [->,redarrow, bend left] node[label, auto=right] {letter to write} (tablet);

    \path (selection) edge [->,redarrow] node[label, auto] {letter/word to write} (laptop);

    \path (-5, 2) edge [->, redarrow] node[label] {ROS} ++(1, 0);
    \path (-5, 2.6) edge [->, greenarrow] node[label] {NaoQI} ++(1, 0);

\caption{\small Overview of the system. In total, the system runs about 10 ROS nodes,
    distributed over the robot itself, a central laptop and Android tablets. Since
    the system does not require any CPU-intensive process, the laptop could be
    removed and the whole logic could run on the robot. Due to the relative
    difficulty to deploy and debug ROS nodes directly on the robot, we however
    relied on the laptop during our experiments.}

%
%\begin{figure}[ht]
%\centering
%
%\resizebox{0.9\linewidth}{!}{%
%
%\begin{tikzpicture}[
%    >=latex,
%    node distance=2cm,
%    every edge/.style={draw, very thick},
%    redarrow/.style={fill=Red, draw=Red},
%    greenarrow/.style={fill=GreenYellow, draw=GreenYellow},
%    yellowarrow/.style={fill=BurntOrange, draw=BurntOrange},
%    cmpt/.style={draw, align=center, rounded corners, inner sep=5pt, font=\sf, fill=black!20},
%    label/.style={midway, align=left, font=\scriptsize\sf, fill=white, above,opacity=0,text opacity=1}]
%
%    \node at (0,0)[cmpt, fill=Cyan!50, text width=2.7cm] (tactile) {tactile interaction \\ manager};
%    \node[cmpt, fill=Cyan!50, left=of tactile] (display) {display \\ manager};
%  
%    \node [cmpt, fill=Green!50, below=of tactile, text width=2.7cm] (shape) {shape position \\ manager};
%    \node [cmpt, fill=Green!50, left=of shape] (learner) {letter \\ learner};
%    \node [cmpt, fill=Green!50, left=of learner] (card) {card \\ detector};
%
%    \node [cmpt, fill=YellowOrange!50, below=1.5cm of shape] (writer) {nao writer};
%    \node [cmpt, fill=YellowOrange!50, left=of writer] (tablet) {tablet detector};
%    \node [cmpt, fill=YellowOrange!50, left=of tablet] (camera) {nao camera};
%    \node [cmpt, fill=YellowOrange!50, below=0.6cm of tablet] (pose) {robot state publisher};
%
%    %%% Relations between components
%    \path (tactile) edge [->] node[label] {clear \\ screen} (display);
%
%    \path ($(shape.north west)!0.33!(shape.north east)$) edge [<-] node[label,left] {clear \\ screen} ($(tactile.south west)!0.33!(tactile.south east)$);
%    \path ($(shape.north west)!0.66!(shape.north east)$) edge [<-, greenarrow] node[label,right] {feedback \\ + location} ($(tactile.south west)!0.66!(tactile.south east)$);
%
%    \path (shape) edge [->, redarrow] node[label, left, align=left] {letter \\to write \\ + time \\ + position} (display);
%    \path (shape) edge [->,redarrow] node[label, right] {letter to write \\ + time \\ + position in\\tablet frame} (writer);
%
%    \path ($(shape.north west)!0.66!(shape.south west)$) edge [->, greenarrow] node[label,below] {feedback \\ + letter} ($(learner.north east)!0.66!(learner.south east)$);
%    \path ($(shape.north west)!0.33!(shape.south west)$) edge [<-, redarrow] node[label] {letter to write \\ + time} ($(learner.north east)!0.33!(learner.south east)$);
%
%    \path (card) edge [->, yellowarrow] node[label] {word to learn} (learner);
%    \path (camera) edge [->, yellowarrow] node[label, right] {camera image} (card);
%    \path (camera) edge [->, redarrow] node[label] {camera image} (tablet);
%    \path (tablet) edge [->, redarrow] node[label] {tablet pose} (writer);
%    \path (pose) edge [->, redarrow] node[label, right] {head pose} (tablet);
%
%    \path (-8, 0) edge [->, redarrow] node[label] {Writing letters} (-7, 0);
%    \path (-8, -0.6) edge [->, greenarrow] node[label] {Responding to feedback} (-7, -0.6);
%    \path (-8, -1.2) edge [->, yellowarrow] node[label] {Detecting card input} (-7, -1.2);
%    
%\end{tikzpicture}
%}
%
%\caption{\small Overview of the system. Components in the top row run on the tablet, 
%those in the middle row on the central controller, and those in the bottom row 
%on the robot.}
%
%    \label{fig:archi}
%\end{figure}
%

@@
